---
title: "615 Midterm Project"
author: "Shicheng Wang"
date: "10/21/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(tidyverse)
library(kableExtra)
```


## Introduction
- **Asia**: filed two-thirds of global patent
- **Us**: residents filed the most overseas patents in 2018
- **Aim**：patent application status/GDP/technology standard.

## Materials description
- **Data**: https://data.worldbank.org (Science technology & Economy growth sector)
- **Access**: WDI package
- **Variables**: GDP/apply/expenditure/export
- **Clean**: NA data/regional values.

```{r,echo=FALSE,include=FALSE}
#install the package
library(WDI)
new_wdi_cache<-WDIcache()

#find out the IDs for the indicators we are intersted in
WDIsearch("Patent applications.")
WDIsearch("Scientific and technical.")
WDIsearch("Research and development expenditure.")
WDIsearch("High-technology exports.")
WDIsearch("gdp.*US\\$", cache = new_wdi_cache)

#download the data
wdi_dat <- WDI(indicator = 
c("IP.PAT.RESD", "IP.PAT.NRES", "GB.XPD.RSDV.GD.ZS","TX.VAL.TECH.CD","NY.GDP.MKTP.CD"), start = 1968, end = 2018, extra = TRUE)
names(wdi_dat)
head(wdi_dat)
```


```{r,echo=FALSE,include=FALSE}
#remove aggregated regional values.
wdi_dat <- subset(wdi_dat, region != "Aggregates")

#rename variables
names(wdi_dat)[which(names(wdi_dat) == "IP.PAT.RESD")] <- "apply_resident" 
names(wdi_dat)[which(names(wdi_dat) == "IP.PAT.NRES")] <- "apply_no_resident" 
names(wdi_dat)[which(names(wdi_dat) == "GB.XPD.RSDV.GD.ZS")] <- "expenditure_gdp"
names(wdi_dat)[which(names(wdi_dat) == "TX.VAL.TECH.CD")] <- "export" 
names(wdi_dat)[which(names(wdi_dat) == "NY.GDP.MKTP.CD")] <- "GDP" 

#clean NA data
da = na.omit(wdi_dat)
dim(da)
```

 
## World average table
```{r,echo=FALSE,include=TRUE}
#process data
da['apply'] = da$apply_resident+da$apply_no_resident
growth <- da %>% group_by(year) %>% 
  summarize(apply=mean(apply),expenditure_gdp=mean(expenditure_gdp),
            export=mean(export),GDP=mean(GDP))

#variables transformation
growth$year <- growth$year
growth$apply <- growth$apply/1000
growth$expenditure_gdp <- growth$expenditure_gdp*10
growth$export <- growth$export*10
growth$GDP <- growth$GDP/100000000000
growth_new <- growth[1:6,]

#produce a table
kable(growth_new, digits = 4, align = "c",booktabs=TRUE ,
caption = "Growth worldwide",col.names = c("year","apply", "expentiture","export","GDP" )) %>% kable_styling(latex_options = "HOLD_position")
```
  

## Variables trend
```{r,echo=FALSE,include=TRUE}
#patent applications trend
p1 = ggplot(growth, aes(x=year, y=apply)) + geom_line(colour='#6600FF') +geom_point(size=4,colour='#6699CC',shape=15)

#expentiture trend
p2 = ggplot(growth, aes(x=year, y=expenditure_gdp)) + geom_line(colour='#660000') +geom_point(size=4,colour='#336600',shape=15)
  
#export trend
p3 = ggplot(growth, aes(x=year, y=export)) + geom_line(colour='#336600') +geom_point(size=4,colour='#333300',shape=15)

#GDP trend
p4 = ggplot(growth, aes(x=year, y=GDP)) + geom_line(colour='#990000')+
  geom_point(size=4,colour='#FFFF99',shape=15)


gridExtra::grid.arrange(p1,p2,p3,p4,ncol = 2)
```


## GDP effect over years 
```{r,echo=FALSE,include=TRUE, warning=FALSE}
income <- da %>% group_by(income,year) %>% summarize(apply=mean(apply),expenditure_gdp=mean(expenditure_gdp),export=mean(export),GDP=mean(GDP))
# income and apply
p5 <- ggplot(data=income, aes(x=year, y=apply,fill=income)) +
      geom_col(stat="identity",position = "dodge") + scale_x_continuous("year",breaks = seq(2007,2017,1))
# income and research and development expenditure (% of GDP)
p6 <- ggplot(data=income, aes(x=year, y=expenditure_gdp,fill=income)) +
      geom_col(stat="identity",position = "dodge") + scale_x_continuous("year",breaks = seq(2007,2017,1))
# income and high-technology exports (current US$)
p7 <- ggplot(data=income, aes(x=year, y=export,fill=income)) +
      geom_col(stat="identity",position = "dodge") + scale_x_continuous("year",breaks = seq(2007,2017,1))

gridExtra::grid.arrange(p5,p6,p7,ncol = 1)
```


## Relationships among variables
```{r,echo=FALSE,include=TRUE}
#group and choose country
country <- da %>% group_by(country) %>%  summarize(apply=mean(apply),expenditure_gdp=mean(expenditure_gdp),export=mean(export),GDP=mean(GDP))
#screening data by GDP
country <- country[order(country$GDP,decreasing=F),]
#variables transformation
country$apply <- 10*log(country$apply)
country$expenditure_gdp <- country$expenditure_gdp*100
country$export <- country$export/100000000
country$GDP <- country$GDP/1000000000

#GDP & apply
p7 <- ggplot(country, aes(x = GDP, y = apply, color = country)) + 
geom_point() +theme(legend.position = 'none')
#GDP & expenditure
p8 <- ggplot(country, aes(x = GDP, y = expenditure_gdp, color = country)) + 
geom_point() +theme(legend.position = 'none')
#GDP & export
p9 <- ggplot(country, aes(x = GDP, y = export, color = country)) + 
geom_point() +theme(legend.position = 'none')
#apply & expenditure
p10 <- ggplot(country, aes(x = apply, y = expenditure_gdp, color = country)) + 
geom_point() +theme(legend.position = 'none')
#apply & export
p11 <- ggplot(country, aes(x = apply, y = export, color = country)) + 
geom_point() +theme(legend.position = 'none')
#expenditure & export
p12 <- ggplot(country, aes(x = expenditure_gdp, y = export, color = country)) + 
geom_point() +theme(legend.position = 'none')

gridExtra::grid.arrange(p7,p8,p9,p10,p11,p12,ncol = 3)
```


## GDP ranking of countries   
```{r,echo=FALSE,include=FALSE}
#group and choose country
country <- da %>% group_by(country) %>%  summarize(apply=mean(apply),expenditure_gdp=mean(expenditure_gdp),export=mean(export),GDP=mean(GDP))

#screening data by GDP
country <- country[order(country$GDP,decreasing=F),]

#variables transformation
country$apply <- log(country$apply)
country$expenditure_gdp <- country$expenditure_gdp*10
country$export <- country$export/10000000000
country$GDP <- country$GDP/100000000000

#produce GDP pieplot
country['sumGDP'] <- cumsum(country$GDP)
country$sumGDP > sum(country$GDP)/2    
country[95,c("country","GDP")] <- c("Other",sum(country[1:89,"GDP"]))
```

```{r,echo=FALSE,include=TRUE}
a <- c(164.6,81.5,51.8,35.9,27.4,317.4)
b <- c("United States","China","Japan","Germany","United Kingdom","Other")
piepercent<- paste(round(100*a/sum(a), 2), "%")

pie(a,labels=piepercent,col= c("skyblue","lightgreen","pink","lavender","lightyellow","lightgrey"),
    radius = 0.85,border="brown",lty=6,main='GDP distribution worldwide')

legend("topright",b,cex=0.5,fill=c("skyblue","lightgreen","pink","lavender","lightyellow","lightgrey"))
```
 

## US data
```{r,echo=FALSE,include=TRUE}
#creat a new dataframe
x <- filter(da, country == "United States") %>% arrange(year)
new <- x[,c(3,6,7,8,16)]

# transformation
new$year <- new$year
new$apply <- new$apply/1000
new$expenditure_gdp <- new$expenditure_gdp*100
new$export <- new$export/1000000000
new$GDP <- new$GDP/10000000000000

new_1 <- new[1:6,]

#US data chart
kable(new_1, digits = 4, align = "c",booktabs=TRUE ,
caption = "Us 10-year data",col.names = c("year", "expentiture","export","GDP","apply" )) %>% kable_styling(latex_options = "HOLD_position")
```


## US's relationship through a decade
```{r}
#GDP & apply
u1 <- ggplot(new, aes(x = GDP, y = apply, color = year)) + 
geom_text(aes(label = year), size=3, nudge_y = 0.2,colour='#660000') +
geom_point(colour='Black',shape=17) 
#GDP & expenditure
u2 <- ggplot(new, aes(x = GDP, y = expenditure_gdp, color = year)) + 
geom_text(aes(label = year), size=3, nudge_y = 0.2,colour='#CC6600') +
geom_point(colour='Black',shape=17) 
#GDP & export
u3 <- ggplot(new, aes(x = GDP, y = export, color = year)) + 
geom_text(aes(label = year), size=3, nudge_y = 0.2,colour='#66CCFF') +
geom_point(colour='Black',shape=17) 
#apply & expenditure
u4 <- ggplot(new, aes(x = apply, y = expenditure_gdp, color = year)) + 
geom_text(aes(label = year), size=3, nudge_y = 0.2,colour='#6666FF') +
geom_point(colour='Black',shape=17) 
#apply & export
u5 <- ggplot(new, aes(x = apply, y = export, color = year)) + 
geom_text(aes(label = year), size=3, nudge_y = 0.2,colour='#336600') +
geom_point(colour='Black',shape=17) 
#expenditure & export
u6 <- ggplot(new, aes(x = expenditure_gdp, y = export, color = year)) + 
geom_text(aes(label = year), size=3, nudge_y = 0.2,colour='#330033') +
geom_point(colour='Black',shape=17) 

gridExtra::grid.arrange(u1,u2,u3,u4,u5,u6,ncol = 2)
```


## Conclusion
- **Five surveys**(worldwide/US)
- **GDP impact**
- **Reality**




